---
title: "Customizing the appearance of your workflowr website"
author: "John Blischak"
date: "2020-06-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

After you've created your workflowr website, you may wish to customize its
appearance. This lesson covers a few key methods. For the full details, please
read the official vignette [Customize your research
website][wflow-02-customization].

[wflow-02-customization]: https://jdblischak.github.io/workflowr/articles/wflow-02-customization.html

## Change the theme

The website you created with workflowr is an [rmarkdown-generated website][rmd-website].
The rmarkdown website settings are saved in the configuration file
`analysis/_site.yml`. The default theme applied to workflowr websites is
"cosmo". You should see the output section below that passes site-wide settings
to `workflowr::wflow_html()`, which in turn passes these arguments to
`rmarkdown::html_document()`.

```
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    theme: cosmo
    highlight: textmate
```

[rmd-website]: https://bookdown.org/yihui/rmarkdown/rmarkdown-site.html

The available themes are from [Bootswatch][], but only a [subset are
available][rmd-themes] for use with rmarkdown:

* cerulean
* cosmo
* darkly
* flatly
* journal
* lumen
* paper
* readable
* sandstone
* simplex
* spacelab
* united
* yeti

[bootswatch]: https://bootswatch.com/
[rmd-themes]: https://bookdown.org/yihui/rmarkdown/html-document.html#appearance-and-style

As an example, this workflowr website uses the theme "
`r yaml::read_yaml("analysis/_site.yml")[["output"]][["workflowr::wflow_html"]][["theme"]]`
".

Go to [Bootswatch][] and find a few themes you like. Then try them out by
changing the value of theme in `analysis/_site.yml` and rebuilding one of the
pages. You can either click the Knit HTML button or use `wflow_build()`. For
convenience, choose a fast running Rmd so you can quickly explore multiple
themes. Note that the updated theme will only be applied to rebuilt pages, so
you'll see the previous theme on different pages if you click around the
website.

Once you've decided on a theme, commit the change to `analysis/_site.yml` while
additional republishing all the Rmd files with the following command:

```{r republish-theme, eval=FALSE}
wflow_publish("analysis/_site.yml", "Change the theme", republish = TRUE)
```

## Change the appearance of the workflowr buttons

You don't have to limit yourself to the pre-made themes. If you know CSS or are
willing to tinker, you can fully customize every aspect of the HTML output.
You do this by [specifying a custom CSS file][custom-css]. In the context of a
workflowr project, you can create the file `analysis/style.css` and then add it
to the `output` section of `analysis/_site.yml`. You can of course name the CSS
whatever you like. The important thing is that you save it in `analysis/` and
specify the path relative to `_site.yml`.

```
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    theme: cosmo
    highlight: textmate
    css: style.css
```

[custom-css]: https://bookdown.org/yihui/rmarkdown/html-document.html#custom-css

The workflowr buttons can be targeted with [CSS class selectors][mdn-class-selectors].
For example, to decrease the font size of the text in all the workflowr buttons,
you could add the following rule:

```
.btn-workflowr {
  font-size: 75%;
}
```

[mdn-class-selectors]: https://developer.mozilla.org/en-US/docs/Web/CSS/Class_selectors

You can also apply styles that are limited to specific types of workflowr
buttons by using the corresponding class:

CSS class  | workflowr button
------------- | -------------
`.btn-workflowr-report` | The button for the workflowr report at the top of the page
`.btn-workflowr-fig`  | The buttons for table of past versions below each figure
`.btn-workflowr-sessioninfo` | The button for the session information at the bottom of the page

For example, to center the buttons of past versions below each figure, you could
add the following rule:

```
.btn-workflowr-fig {
  display: block;
  margin: auto;
}
```

Similar to above, once you are satisfied with the changes you've made to the
appearance, you can commit the changes and republish the webpages:

```{r republish-css, eval=FALSE}
wflow_publish(c("analysis/_site.yml", "analysis/style.css"),
              "Style the workflowr buttons",
              republish = TRUE)
```
